#libraries
```{r}
      library(GISTools)
      library(sf)
      library(sp)
      library(tmap)
      library(rgdal)
      library(tidyverse)
      library(ggplot2)
      library(gridExtra)
      library(grid)
```

#objects
```{r}
      acs<-st_read("C:\\Users\\tayqu\\My Drive\\School\\01 Current Semester\\R\\Assignments\\assignment 1\\acs.shp")

      acs_sf = st_as_sf(acs)
      
      roads<-st_read("C:\\Users\\tayqu\\My Drive\\School\\01 Current Semester\\R\\Assignments\\assignment 1\\roads.shp")
      
      roads_sf = st_as_sf(roads)
```

#question1
```{r}
      colnames(acs)
      head(acs)
      nrow(acs)
      
      colnames(roads)
      head(roads)
      nrow(roads)
```

  #total population
```{r}
      mean(acs$pop_ttl)
      median(acs$pop_ttl)
      min(acs$pop_ttl)
      max(acs$pop_ttl)
```
  
  #percent population with a college degree
```{r}
      mean(is.na(acs$college))
      median(acs$college,na.rm=T)
      min(is.na(acs$college))
      max(is.na(acs$college))
```
  
  #median household income
```{r}
      mean(is.na(acs$income))
      median(acs$income,na.rm=T)
      min(is.na(acs$income))
      max(is.na(acs$income))
```

  #poverty rate (poverty)
```{r}
      mean(is.na(acs$poverty))
      median(acs$poverty,na.rm=T)
      min(is.na(acs$poverty))
      max(is.na(acs$poverty))
```
  
  #percent population without health insurance (n_nsrnc)
```{r}
      mean(is.na(acs$n_nsrnc))
      median(acs$n_nsrnc,na.rm=T)
      min(is.na(acs$n_nsrnc))
      max(is.na(acs$n_nsrnc))
```
  
  #percent households without a vehicle (n_vhcls)
```{r}
      mean(is.na(acs$n_vhcls))
      median(acs$n_vhcls,na.rm=T)
      min(is.na(acs$n_vhcls))
      max(is.na(acs$n_vhcls))
```
  
#question2
```{r}
      plot1<-ggplot(acs, aes(x = "",pop_ttl)) + geom_boxplot()
      plot2<-ggplot(acs, aes(x = "",college)) + geom_boxplot()
      plot3<-ggplot(acs, aes(x = "",income)) + geom_boxplot()
      plot4<-ggplot(acs, aes(x = "",poverty)) + geom_boxplot()
      plot5<-ggplot(acs, aes(x = "",n_nsrnc)) + geom_boxplot()
      plot6<-ggplot(acs, aes(x = "",n_vhcls)) + geom_boxplot()
      
      grid.arrange(plot1,plot2,plot3,plot4,plot5,plot6)
```
  #graph with trend line
```{r}
        ggplot(data = acs, mapping = aes(x = income, y = n_nsrnc)) + geom_point() + geom_smooth(method = "lm")
```
#Saving scatterplot
```{r}
png(file="C:\\Users\\tayqu\\My Drive\\School\\01 Current Semester\\R\\Assignments\\assignment 1\\scatterplot.png")
      
      ggplot(data = acs, mapping = aes(x = income, y = n_nsrnc)) + geom_point() + geom_smooth(method = "lm")

      
      dev.off()
```

#question3
  
  #percent population with a college degree
```{r}
       p1<-tm_shape(acs_sf) +
          tm_polygons("college", title = "% population with college degree", palette = "Blues", 
                    legend.hist = T) +
          tm_scale_bar(width = 0.2) +
          tm_compass(position = c(.8,.7)) +
          tm_layout(frame = F, legend.hist.size = 0.5)
```
  
  #median household income
```{r}
      p2<-tm_shape(acs_sf) +
          tm_polygons("income", title = "median household income", palette = "Greens", 
                    legend.hist = T) +
          tm_scale_bar(width = 0.2) +
          tm_compass(position = c(.8,.7)) +
          tm_layout(frame = F, legend.hist.size = 0.5)
```
  
   #percent population without health insurance
```{r}
      p3<-tm_shape(acs_sf) +
          tm_polygons("n_nsrnc", title = "% population without health insurance", palette = "Reds", 
                    legend.hist = T) +
          tm_scale_bar(width = 0.2) +
          tm_compass(position = c(.8,.7)) +
          tm_layout(frame = F, legend.hist.size = 0.5)
```
     
  #percent households without a vehicle
```{r}
      p4<-tm_shape(acs_sf) +
          tm_polygons("n_vhcls", title = "% households without a vehicle", palette = "Purples", 
                    legend.hist = T) +
          tm_scale_bar(width = 0.2) +
          tm_compass(position = c(.8,.7)) +
          tm_layout(frame = F, legend.hist.size = 0.5) 
```
  
  #map grid
```{r}
      tmap_arrange(p1,p2,p3,p4)

```
  
  #saving map
```{r}
      png(file="C:\\Users\\tayqu\\My Drive\\School\\01 Current Semester\\R\\Assignments\\assignment 1\\acsmap.png")
      
      tmap_arrange(p1,p2,p3,p4)
      
      dev.off()
```
  
#question4
```{r}
      f<-st_union(acs_sf)

      roadmap<-tm_shape(roads)+
               tm_lines("RTTYP", lwd = 1, title.col = "Road Type", palette = "Spectral")+
               tm_shape(f)+
               tm_borders(lwd = 3)
      
      roadmap
```

  #Saving roadmap
```{r}
      png(file="C:\\Users\\tayqu\\My Drive\\School\\01 Current Semester\\R\\Assignments\\assignment 1\\roadmap.png")
      
      roadmap
      
      dev.off()
```